PANCHANGE_data_preprocessing

### 2020-08-14 PANCHANGE_data_preprocessing2020-07-15.Rmd
PHQ section adapted for NBR, confirmed it works.

```{r coping pcl base sum score if missigng items set to NA}
coping.pcl.raw.id <- coping.pcl.raw.id %>%
  mutate(
    pcl.sum_score_base =
      if_else(condition = na_per_person_pcl_base > 0,
              true = NA_real_,
              false = pcl.sum_score_base_raw,
              missing = NA_real_
      )
  )

@@

```{r coping pcl retro sum score if missigng items set to NA}
coping.pcl.raw.id <- coping.pcl.raw.id %>%
  mutate(
    pcl.sum_score_retro =
      if_else(condition = na_per_person_pcl_retro > 0,
              true = NA_real_,
              false = pcl.sum_score_retro_raw,
              missing = NA_real_
      )
  )

@@

Will need to double check these namings once we have the new cleaned data -- KT. 
```{r coping nbr phq cleaning}
if(NBR == TRUE) {
coping.nbr.phq.raw.id <- coping.nbr.phq.raw %>%
  drop_na(subjectid) %>% # Drop NAs
  distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(subjectid, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
  separate(ID.intm, into = "ID", sep = 7) %>%
  mutate(ID = as.numeric(ID)) %>%
  select(
    Sample, # Sample
    ID, # ID
    phq.problems_made_difficult_care,
    phq.pandemic_felt_feelings,
    phq.little_interest_or_pleasure_in_doing_things_retro = phq.little_interest_or_pleasure_in_doing_things,
    phq.feeling_down_depressed_or_hopeless_retro = phq.feeling_down_depressed_or_hopeless,
    phq.staying_asleep_sleeping_trouble_retro = phq.staying_asleep_sleeping_trouble,
    phq.feeling_tired_or_having_little_energy_retro = phq.feeling_tired_or_having_little_energy,
    phq.poor_appetite_or_overeating_retro = phq.poor_appetite_or_overeating,
    phq.feeling_bad_failure_family_retro = phq.feeling_bad_failure_family,
    phq.trouble_concentrating_reading_newspaper_retro = phq.trouble_concentrating_newspaper_reading, # differ NBR prepandemic
    phq.moving_fidgety_noticed_opposite_retro = phq.moving_fidgety_opposite_slowly, # differ NBR prepandemic
    phq.dead_hurting_thoughts_retro = phq.dead_hurting_thoughts,
    phq.little_interest_or_pleasure_in_doing_things_base = phq.little_interest_or_pleasure_in_doing_things.1,
    phq.feeling_down_depressed_or_hopeless_base = phq.feeling_down_depressed_or_hopeless.1,
    phq.staying_asleep_sleeping_trouble_base = phq.staying_asleep_sleeping_trouble.1,
    phq.feeling_tired_or_having_little_energy_base = phq.feeling_tired_or_having_little_energy.1,
    phq.poor_appetite_or_overeating_base = phq.poor_appetite_or_overeating.1,
    phq.feeling_bad_failure_family_base = phq.feeling_bad_failure_family.1,
    phq.trouble_concentrating_reading_newspaper_base = phq.trouble_concentrating_newspaper_reading.1, # differ NBR prepandemic
    phq.moving_fidgety_noticed_opposite_base = phq.moving_fidgety_opposite_slowly.1, # differ NBR prepandemic
    phq.dead_hurting_thoughts_base = phq.dead_hurting_thoughts.1
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r coping phq base sum score if missigng items set to NA}
coping.phq.raw.id <- coping.phq.raw.id %>%
  mutate(
    phq.sum_score_base =
      if_else(condition = na_per_person_phq_base > 0,
              true = NA_real_,
              false = phq.sum_score_base_raw,
              missing = NA_real_
      )
  )

@@

8 items
```{r phq base sum score if missigng items set to NA - 8 items}
coping.phq.raw.id <- coping.phq.raw.id %>%
  mutate(
    phq.sum_score_8items_base =
      if_else(condition = na_per_person_phq_base > 1 &
                phq.missing_only_suicide_item_base == 1,
              true = NA_real_,
              false = phq.sum_score_base_raw,
              missing = NA_real_
      )
  )

@@

```{r coping phq retro sum score if missigng items set to NA}
coping.phq.raw.id <- coping.phq.raw.id %>%
  mutate(
    phq.sum_score_retro =
      if_else(condition = na_per_person_phq_retro > 0,
              true = NA_real_,
              false = phq.sum_score_retro_raw,
              missing = NA_real_
      )
  )

@@

8 items
```{r phq retro sum score if missigng items set to NA - 8 items}
coping.phq.raw.id <- coping.phq.raw.id %>%
  mutate(
    phq.sum_score_8items_retro =
      if_else(condition = na_per_person_phq_retro > 1 &
                phq.missing_only_suicide_item_retro == 1,
              true = NA_real_,
              false = phq.sum_score_retro_raw,
              missing = NA_real_
      )
  )

>>>>

```{r coping pcl base sum score if missing items set to NA}
coping.pcl.raw.id <- coping.pcl.raw.id %>%
  mutate(
    pcl.sum_score_base =
      if_else(condition = na_per_person_pcl_base > 0,
              true = NA_real_,
              false = pcl.sum_score_base_raw,
              missing = NA_real_
      )
  )

@@

```{r coping pcl retro sum score if missing items set to NA}
coping.pcl.raw.id <- coping.pcl.raw.id %>%
  mutate(
    pcl.sum_score_retro =
      if_else(condition = na_per_person_pcl_retro > 0,
              true = NA_real_,
              false = pcl.sum_score_retro_raw,
              missing = NA_real_
      )
  )

@@

Will need to double check these namings once we have the new cleaned data -- KT. 
```{r coping nbr phq cleaning}
if(NBR == TRUE) {
coping.nbr.phq.raw.id <- coping.nbr.phq.raw %>%
  drop_na(subjectid) %>% # Drop NAs
  distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(subjectid, into = c("Sample", "ID.intm"), sep = 6) %>% # Split ID in Sample and Number
  separate(ID.intm, into = "ID", sep = 7) %>%
  mutate(ID = as.numeric(ID)) %>%
  select(
    Sample, # Sample
    ID, # ID
    phq.problems_made_difficult_care = phq.problems_made_care_difficult,
    phq.pandemic_felt_feelings,
    phq.little_interest_or_pleasure_in_doing_things_retro = phq.little_interest_or_pleasure_in_doing_things,
    phq.feeling_down_depressed_or_hopeless_retro = phq.feeling_down_depressed_or_hopeless,
    phq.staying_asleep_sleeping_trouble_retro = phq.staying_asleep_sleeping_trouble,
    phq.feeling_tired_or_having_little_energy_retro = phq.feeling_tired_or_having_little_energy,
    phq.poor_appetite_or_overeating_retro = phq.poor_appetite_or_overeating,
    phq.feeling_bad_failure_family_retro = phq.family_feeling_bad_failure,
    phq.trouble_concentrating_reading_newspaper_retro = phq.trouble_concentrating_newspaper_reading, # differ NBR prepandemic
    phq.moving_fidgety_noticed_opposite_retro = phq.moving_fidgety_opposite_slowly, # differ NBR prepandemic
    phq.dead_hurting_thoughts_retro = phq.dead_hurting_thoughts,
    phq.little_interest_or_pleasure_in_doing_things_base = phq.little_interest_or_pleasure_in_doing_things.1,
    phq.feeling_down_depressed_or_hopeless_base = phq.feeling_down_depressed_or_hopeless.1,
    phq.staying_asleep_sleeping_trouble_base = phq.staying_asleep_sleeping_trouble.1,
    phq.feeling_tired_or_having_little_energy_base = phq.feeling_tired_or_having_little_energy.1,
    phq.poor_appetite_or_overeating_base = phq.poor_appetite_or_overeating.1,
    phq.feeling_bad_failure_family_base = phq.family_feeling_bad_failure.1,
    phq.trouble_concentrating_reading_newspaper_base = phq.trouble_concentrating_newspaper_reading.1, # differ NBR prepandemic
    phq.moving_fidgety_noticed_opposite_base = phq.moving_fidgety_opposite_slowly.1, # differ NBR prepandemic
    phq.dead_hurting_thoughts_base = phq.dead_hurting_thoughts.1
  ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

@@

```{r coping phq base sum score if missing items set to NA}
coping.phq.raw.id <- coping.phq.raw.id %>%
  mutate(
    phq.sum_score_base =
      if_else(condition = na_per_person_phq_base > 0,
              true = NA_real_,
              false = phq.sum_score_base_raw,
              missing = NA_real_
      )
  )

@@

8 items
```{r phq base sum score if missing items set to NA - 8 items}
coping.phq.raw.id <- coping.phq.raw.id %>%
  mutate(
    phq.sum_score_8items_base =
      if_else(condition = na_per_person_phq_base > 1 &
                phq.missing_only_suicide_item_base == 1,
              true = NA_real_,
              false = phq.sum_score_base_raw,
              missing = NA_real_
      )
  )

@@

```{r coping phq retro sum score if missing items set to NA}
coping.phq.raw.id <- coping.phq.raw.id %>%
  mutate(
    phq.sum_score_retro =
      if_else(condition = na_per_person_phq_retro > 0,
              true = NA_real_,
              false = phq.sum_score_retro_raw,
              missing = NA_real_
      )
  )

@@

8 items
```{r phq retro sum score if missing items set to NA - 8 items}
coping.phq.raw.id <- coping.phq.raw.id %>%
  mutate(
    phq.sum_score_8items_retro =
      if_else(condition = na_per_person_phq_retro > 1 &
                phq.missing_only_suicide_item_retro == 1,
              true = NA_real_,
              false = phq.sum_score_retro_raw,
              missing = NA_real_
      )
  )
